---
title: "Bike Share"
output:
  html_document: default
  pdf_document: default
date: "2024-02-23"
---

###Install Packages
```{r}
library(tidyverse)
library(conflicted)
```

###Combined all 3 months and view
tripdata_jan23_mar23 <- rbind(X202301_divvy_tripdata, X202302_divvy_tripdata, X202303_divvy_tripdata)


###View combined data

View(tripdata_jan23_mar23)

###View column names
colnames(tripdata_jan23_mar23)


###Remove columns we wont use
tripdata_jan23_mar23 <- tripdata_jan23_mar23 %>%
  select(-c(start_station_name, start_station_id, end_station_name, end_station_id, start_lat, start_lng, end_lat, end_lng,))

###View data 

View(tripdata_jan23_mar23)

###Review new table
```{r}
colnames(tripdata_jan23_mar23)
nrow(tripdata_jan23_mar23)
head(tripdata_jan23_mar23)
str(tripdata_jan23_mar23)
```
###New colums for date, month, day and year
```{r}
tripdata_jan23_mar23$date <- as.Date(tripdata_jan23_mar23$started_at) 
tripdata_jan23_mar23$month <- format(as.Date(tripdata_jan23_mar23$date), "%m")
tripdata_jan23_mar23$day <- format(as.Date(tripdata_jan23_mar23$date), "%d")
tripdata_jan23_mar23$year <- format(as.Date(tripdata_jan23_mar23$date), "%Y")
tripdata_jan23_mar23$day_of_week <- format(as.Date(tripdata_jan23_mar23$date), "%A")
```
###Add ride_length
```{r}
tripdata_jan23_mar23$ride_length <- difftime(tripdata_jan23_mar23$ended_at,tripdata_jan23_mar23$started_at)
```

###View data types
```{r}
str(tripdata_jan23_mar23)
```
###Convert ride_length
```{r}
is.factor(tripdata_jan23_mar23$ride_length)
tripdata_jan23_mar23$ride_length <- as.numeric(as.character(tripdata_jan23_mar23$ride_length))
is.numeric(tripdata_jan23_mar23$ride_length)
```

###Compare members and casual users
```{r}
aggregate(tripdata_jan23_mar23$ride_length ~ tripdata_jan23_mar23$member_casual, FUN = mean)
aggregate(tripdata_jan23_mar23$ride_length ~ tripdata_jan23_mar23$member_casual, FUN = median)
aggregate(tripdata_jan23_mar23$ride_length ~ tripdata_jan23_mar23$member_casual, FUN = max)
aggregate(tripdata_jan23_mar23$ride_length ~ tripdata_jan23_mar23$member_casual, FUN = min)
```
###the average ride time by each day for members vs casual users
```{r}
aggregate(tripdata_jan23_mar23$ride_length ~ tripdata_jan23_mar23$member_casual + tripdata_jan23_mar23$day_of_week, FUN = mean)
```
###Arrange days of week
```{r}
tripdata_jan23_mar23$day_of_week <- ordered(tripdata_jan23_mar23$day_of_week, levels=c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"))
```

###Run again
```{r}
aggregate(tripdata_jan23_mar23$ride_length ~ tripdata_jan23_mar23$member_casual + tripdata_jan23_mar23$day_of_week, FUN = mean)
```
###analyze ridership data by type and weekday
```{r}
tripdata_jan23_mar23 %>% 
  mutate(weekday = wday(started_at, label = TRUE)) %>%  
  group_by(member_casual, weekday) %>%
  summarise(number_of_rides = n()							
            ,average_duration = mean(ride_length)) %>% 		
  arrange(member_casual, weekday)	
```
###Visualize the number of rides by rider type
```{r}
tripdata_jan23_mar23 %>% 
  mutate(weekday = wday(started_at, label = TRUE)) %>% 
  group_by(member_casual, weekday) %>% 
  summarise(number_of_rides = n()
            ,average_duration = mean(ride_length)) %>% 
  arrange(member_casual, weekday)  %>% 
  ggplot(aes(x = weekday, y = number_of_rides, fill = member_casual)) +
  geom_col(position = "dodge")
```
###Visualization for average duration
```{r}
tripdata_jan23_mar23 %>% 
  mutate(weekday = wday(started_at, label = TRUE)) %>% 
  group_by(member_casual, weekday) %>% 
  summarise(number_of_rides = n()
            ,average_duration = mean(ride_length)) %>% 
  arrange(member_casual, weekday)  %>% 
  ggplot(aes(x = weekday, y = average_duration, fill = member_casual)) +
  geom_col(position = "dodge")
```

###Bikes rented
```{r}
tripdata_jan23_mar23 %>% 
  ggplot(aes(x = rideable_type, fill = member_casual)) + geom_bar(position = "dodge") +
  labs(x = "Type of bike", y = "Number of rentals", title = "Rented bikes", fill = "Type of membership")
```








